A Wait-free Queue with Polylogarithmic Step Complexity

### 2023-01-09 podc_paper/correctness-2.tex
cleaning up a few things

We first bound the size of RBTs.  Let $q_{max}$ be the maximum size of the queue at any time during the sequential execution given by the linearization $L$.
Recall that GC is done on a node every $G$ times its RBT is updated, and we chose $G$ to be $p^2\ceil{\log p}$.

%We say the index $b$ \emph{is marked in node $v$} if some process has tested
%whether $b>\var{v}.\fld{last}[*]$ at line \ref{marktest1} or \ref{marktest2}.
%Intuitively, a block's index value is marked 
%when a dequeue passes through this block to find the enqueue
%whose argument it should return, or when a dequeue that returns null passes through the block.
%Thus, $b$ gets marked in $v$ when $v.\fld{blocks}[b]$ 
%contains either an enqueue whose value is being dequeued or a \nl\ dequeue.
%If the index of block $B$ is marked in $v$, we also say $B$ is marked in $v$.

@@

\begin{theorem}\label{spaceBound}
The size of our queue data structure is $O(pq_{max}+p^3\log p)$ words of memory.
\end{theorem}
\begin{proof}
There are $2p-1$ nodes in the ordering tree.  Aside from the RBT, each node uses $O(1)$ memory words.
So the space bound follows from \Cref{RBTbound} and the fact that $G$ is chosen
to be $p^2\ceil{\log p}$.
\end{proof}

@@

@@

The idea above leads us to a poly-log data structure that supports throwing away all the blocks with keys smaller than an index. Red-black trees do this for us. \nf{Get(i)}, \nf{Append()} and \nf{Split(i)} are logarithmic in block trees.
We can create a shared red-black tree just creating a new path for the operation and then using \nf{CAS} to change the root of the tree. See [this] for more.

\begin{observation}
PBRT supports poly-log operations ....
\end{observation}

@@

\begin{observation}
Every $p^2$ block appended to the root, \nf{FreeMemory} is invoked.  
\end{observation}

To know the last block dequeied from we can implement a shared array among processes which they write the last root block they have dequeued from.

\begin{lemma}
    $\nf{Max(Last)} - \text{index of the last finished block}$ in the node $n$ is $O(p)$.
\end{lemma}

\begin{lemma}
    After \nf{FreeMemory}, the space taken by each node of the tree is $O(p+q)$.
\end{lemma}

\begin{corollary}
The space taken by each node of the tree is $O(p^2+q)$. The total space in the tree is $PolyLog(p+q)$.
\end{corollary}

\begin{lemma}
  The amortized step per process for the algorithm with garbage collection is $PolyLog(p+q)$.
\end{lemma}

@@

\here{Remark on an optimization:  instead of storing \head\ field separate from \fld{blocks}, we could store
the maximum index of the RBT in the root of the RBT.  This would simplify things, but require modifying the proof more extensively.}

>>>>

We first bound the size of RBTs.  Let $q_{max}$ be the maximum size of the queue at any time during the sequential execution given by the linearization $L$.
Recall that GC is done on a node every $G$ times its RBT is updated, and we chose $G$ to be $p^2\ceil{\log p}$.
Part of the proof of the following lemma is similar to the proof of \Cref{dSearchTime}.
%We say the index $b$ \emph{is marked in node $v$} if some process has tested
%whether $b>\var{v}.\fld{last}[*]$ at line \ref{marktest1} or \ref{marktest2}.
%Intuitively, a block's index value is marked 
%when a dequeue passes through this block to find the enqueue
%whose argument it should return, or when a dequeue that returns null passes through the block.
%Thus, $b$ gets marked in $v$ when $v.\fld{blocks}[b]$ 
%contains either an enqueue whose value is being dequeued or a \nl\ dequeue.
%If the index of block $B$ is marked in $v$, we also say $B$ is marked in $v$.

@@

The following theorem bounds the space that is reachable (and therefore cannot be freed by the environment's garbage collector) at any time.
\here{There may be room to improve this bound by a more careful analysis since the bounds so far
didn't use the fact that operations only appear along one path in the tree; in other words
doing a bound on total size of all RBTs at one level of the ordering tree at once, rather than
on each node separately, might reduce it to qlogp + poly(p).  Might also be possible to be
more precise about how it depends on "current" size of queue rather than max size.}

\begin{theorem}\label{spaceBound}
The queue data structure uses a maximum of $O(pq_{max}+p^3\log p)$ words of memory at any time.
\end{theorem}
\begin{proof}
There are $2p-1$ nodes in the ordering tree.  Aside from the RBT, each node uses $O(1)$ memory words.
Each process may hold pointers to $O(1)$ RBTs that are no longer current in local variables.
So the space bound follows from \Cref{RBTbound} and the fact that $G$ is chosen
to be $p^2\ceil{\log p}$.
\end{proof}

@@

We remark that the implementation remains wait-free:  the depth of recursion in each routine is  bounded
by the height of the tree and there are no loops.

@@

@@

@@
